Removal and Replacement only

### 2020-08-07 analysis/boneyard.Rmd
now with suppl fig 3 and fig 3

>>>>

## tsne

```{r tSNE}

# removal
hyptsne <- subsetmaketsne("hypothalamus", removallevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", removallevels, sexlevels)
gontsne <- subsetmaketsne("gonads", removallevels, sexlevels)

pittsne <- addgroupings(pittsne)

t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
  facet_wrap(~sex) 
t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank())
t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank())
t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
t456

# replacement
hyptsne <- subsetmaketsne("hypothalamus", timinglevels, sexlevels)
pittsne <- subsetmaketsne("pituitary", timinglevels, sexlevels)
gontsne <- subsetmaketsne("gonads", timinglevels, sexlevels)

pittsne <- addgroupings(pittsne)

t4 <- plottsneelipsev2(pittsne, pittsne$treatment, allcolors) +  labs(title = "pituitary ", subtitle = "parental stage and manipulations") +
  facet_wrap(~sex) 
t5 <- plottsneelipsev2(pittsne, pittsne$extint, colorhypothesis) +  labs(title = "  ", subtitle = "external enviornment")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank())
t6 <- plottsneelipsev2(pittsne, pittsne$earlylate, colorhypothesis) +  labs(title = "  ", subtitle = "internal clock")   +
  facet_wrap(~sex) + theme(axis.title.y = element_blank())
t456 <- plot_grid(t4, t5, t6, nrow = 1, rel_widths = c(1.1,1,1))
t456

### 2020-08-07 analysis/figure-3.Rmd
now with suppl fig 3 and fig 3

```{r fig3, fig.height=7.25, fig.width=7.25}
plotcandidatemanp <- function(df, whichgene, whichtissue){

  p <- df %>%
    dplyr::filter(gene  == whichgene)  %>%
    ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
    geom_boxplot(outlier.shape = NA, lwd=0.5) +
    #facet_wrap(~sex) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    theme_B3() +
    theme(legend.position = "none",

          axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(face = "italic")) +
    scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors)  +
    labs(title = whichtissue, 
         subtitle = whichgene, 
         y = "gene expression", x = "Semi-sequential stages") +
    geom_signif(comparisons = list(c("inc.d9", "early"), 
                                  c( "inc.d17", "prolong"), 
                                  c( "prolong", "hatch")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075) +
  geom_signif(comparisons = list(c( "inc.d3", "m.inc.d3"),
                                   c( "inc.d9", "m.inc.d9"),
                                   c( "inc.d17", "m.inc.d17"),
                                   c( "hatch", "m.n2")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0,  size = 0.5, step_increase = 0.075) +
    geom_signif(comparisons = list(c( "hatch", "extend"), 
                                  c( "n5", "extend")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075)

@@

  p1 <- df %>%
    dplyr::filter(gene  == whichgene)  %>%
    mutate(earlylate = factor(earlylate, levels = levelhypothesis)) %>%
    ggplot(aes(y =  counts, x = earlylate, fill = earlylate, color = sex)) +
    geom_boxplot(outlier.shape = NA, lwd=0.5) +
    #facet_wrap(~sex) +
    geom_jitter(size = 0.25, aes(color = treatment)) +
    theme_B3() +
  scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors)  +
  geom_signif(comparisons = list(c( "reference", "early"), 
                                  c( "early", "late"),
                                 c( "reference", "late")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075) +
  labs(title = paste(whichtissue, sep = " "), subtitle = " ", y = NULL) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- df %>%
    dplyr::filter(gene  == whichgene)  %>%
      mutate(extint = factor(extint, levels = levelhypothesis)) %>%
    ggplot(aes(y =  counts, x = extint, fill = extint, color = sex)) +
    geom_boxplot(outlier.shape = NA, lwd=0.5) +
    #facet_wrap(~sex) +
    geom_jitter(size = 0.25, aes(color = treatment)) +
    theme_B3() +
  scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors)  +
  geom_signif(comparisons = list(c( "reference", "eggs"), 
                                  c( "eggs", "chicks"),
                                 c( "chicks", "loss"),
                                 c( "reference", "chicks"),
                                 c( "eggs", "loss"),
                                 c( "reference", "loss")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075) +
  labs(title = " ", subtitle = " " , y = NULL) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

p <- plot_grid(p1,p2)
return(p)
}

a <- plotcandidatemanp(hypvsdf, "HTR2C", "female hypothalamus") 
b <- boxplotextint(hypvsdf, "HTR2C", " ") 
c <- plotcandidatemanp(hypvsdm, "HTR2C", "male hypothalamus") 
d <- boxplotextint(hypvsdm, "HTR2C", " ")

e <- plotcandidatemanp(pitvsdf, "PRL", "female pituitary")
f <- boxplotextint(pitvsdf, "PRL", " ")

g <- plotcandidatemanp(pitvsdm, "PRL", "male pituitary")
h <- boxplotextint(pitvsdm, "PRL", " ")

i <- plotcandidatemanp(gonvsdf, "ESR2", "female gonads") 
j <- boxplotextint(gonvsdf, "ESR2", " ")

k <- plotcandidatemanp(gonvsdm, "ESR2", "male gonads") 
l <- boxplotextint(gonvsdm, "ESR2", " ")

fig3 <- plot_grid(a,b,c,d,e,f, 
                       g,h,i,j,k,l, rel_widths = c(1.75,1,1.75,1),
                       labels = c("A", "", "", "", 
                                  "B", "", "", "", 
                                  "C", "", "", ""),
                       label_size = 8, hjust = 0)
fig3
```

## write files

@@

>>>>

```{r }
plotcandidatemanp <- function(df, whichgene, whichtissue){

  p <- df %>%
    dplyr::filter(gene  == whichgene)  %>%
    ggplot(aes(y =  counts, x = treatment, fill = treatment, color = sex)) +
    geom_boxplot(outlier.shape = NA, lwd=0.5) +
    #facet_wrap(~sex) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    theme_B3() +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1),
          plot.subtitle = element_text(face = "italic")) +
    scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors)  +
    labs( subtitle = whichgene, 
         y = "gene expression", x = "Semi-sequential stages") +
    geom_signif(comparisons = list(c("inc.d9", "early"), 
                                  c( "inc.d17", "prolong"), 
                                  c( "prolong", "hatch")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075) +
  geom_signif(comparisons = list(c( "inc.d3", "m.inc.d3"),
                                   c( "inc.d9", "m.inc.d9"),
                                   c( "inc.d17", "m.inc.d17"),
                                   c( "hatch", "m.n2")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0,  size = 0.5, step_increase = 0.075) +
    geom_signif(comparisons = list(c( "hatch", "extend"), 
                                  c( "n5", "extend")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075)

@@

  p1 <- df %>%
    dplyr::filter(gene  == whichgene)  %>%
    mutate(earlylate = factor(earlylate, levels = levelhypothesis)) %>%
    ggplot(aes(y =  counts, x = earlylate, fill = earlylate, color = sex)) +
    geom_boxplot(outlier.shape = NA, lwd=0.5) +
    #facet_wrap(~sex) +
    geom_jitter(size = 0.25, aes(color = treatment)) +
    theme_B3() +
  scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors)  +
  geom_signif(comparisons = list(c( "reference", "early"), 
                                  c( "early", "late"),
                                 c( "reference", "late")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075) +
  labs(subtitle = " ", y = NULL) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- df %>%
    dplyr::filter(gene  == whichgene)  %>%
      mutate(extint = factor(extint, levels = levelhypothesis)) %>%
    ggplot(aes(y =  counts, x = extint, fill = extint, color = sex)) +
    geom_boxplot(outlier.shape = NA, lwd=0.5) +
    #facet_wrap(~sex) +
    geom_jitter(size = 0.25, aes(color = treatment)) +
    theme_B3() +
  scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors)  +
  geom_signif(comparisons = list(c( "reference", "eggs"), 
                                  c( "eggs", "chicks"),
                                 c( "chicks", "loss"),
                                 c( "reference", "chicks"),
                                 c( "eggs", "loss"),
                                 c( "reference", "loss")),
                map_signif_level=TRUE,
                textsize = 2, family = 'Helvetica',
                vjust = 0, size = 0.5, step_increase = 0.075) +
  labs(subtitle = " " , y = NULL) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

p <- plot_grid(p1,p2, rel_widths = c(3.5,4))
return(p)
}

## DEGs

```{r allDEGs}
DEG_path <- "../results/DEseq2/treatment/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
#DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) # %>%
 # mutate(tissue = factor(tissue, levels = tissuelevel),
  #       comparison = factor(comparison , levels = comparisonlevelsremoval),
  #       direction = factor(direction, levels = alllevels))
head(allDEG)

candidateDEGS <- allDEG %>%
  filter(gene %in% parentalcaregenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", 
                         "gonad" = "G", "gonads" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, inc.d3_m.inc.d3, inc.d9_m.inc.d9, inc.d17_m.inc.d17, hatch_m.n2,
         inc.d17_prolong, hatch_prolong,extend_hatch, n5_extend, 
         early_inc.d9 ,hatch_early) %>%
  arrange(gene)
head(candidateDEGS[1:5])

DEGmanip <- allDEG %>% 
  filter(comparison %in% comparisonlevelsmanip) %>%
  mutate(comparison = factor(comparison, levels = comparisonlevelsmanip))

DEGmanipf <- DEGmanip %>% filter(sex == "female")
DEGmanipm <- DEGmanip %>% filter(sex == "male")

DEGmanip %>% 
  group_by(comparison, sex, tissue) %>%
  summarize(totalDEGs = n())  %>%
  arrange(desc(totalDEGs))
```

## fig

```{r fig3, fig.height=7.25, fig.width=7.25}
a <- plotcandidatemanp(hypvsdf, "HTR2C", "female hypothalamus") 
b <- boxplotextint(hypvsdf, "HTR2C", " ")

d <- plotcandidatemanp(pitvsdf, "PRL", "female pituitary")
e <- boxplotextint(pitvsdf, "PRL", " ")

g <- plotcandidatemanp(gonvsdf, "ESR2", "female gonads") 
h <- boxplotextint(gonvsdf, "ESR2", " ")

c <- makebargraph(DEGmanipf, "hypothalamus","DEGs w/ + LFC", 0, 3200, DEGmanipf$comparison) + 
  labs(title = "female hypothalamus", x = "Comparison of manipulated group to parental controls") + 
  theme(strip.text = element_blank()) + 
  scale_x_discrete(breaks= comparisonlevelsmanip,
                      labels= comparisonlabelssmanip,
                   drop=FALSE)

f <- makebargraph(DEGmanipf, "pituitary","DEGs w/ + LFC", 0, 3200, DEGmanipf$comparison) + 
  labs(title = "female pituitary",x = "Comparison of manipulated group to parental controls")  + 
  theme(strip.text = element_blank()) + 
  scale_x_discrete(breaks= comparisonlevelsmanip,
                      labels= comparisonlabelssmanip,
                   drop=FALSE)

i <- makebargraph(DEGmanipf, "gonad","DEGs w/ + LFC", 0, 3200, DEGmanipf$comparison) + 
  theme(strip.text = element_blank()) + 
  labs(title = "female hypothalamus",x = "Comparison of manipulated group to parental controls") + 
  scale_x_discrete(breaks= comparisonlevelsmanip,
                      labels= comparisonlabelssmanip,
                   drop=FALSE)
fig3 <- plot_grid(c, a,b,f,d,e,i,g,h, ncol = 3, rel_widths = c(2,1.8,1), 
                  labels = c("A", "", "",
                             "B", "", "",
                             "C", "", ""), label_size = 8, hjust = 0)
fig3

## suppl fig

```{r supplfig-3}

a <- plotcandidatemanp(hypvsdm, "HTR2C", "male hypothalamus") 
b <- boxplotextint(hypvsdm, "HTR2C", " ")

d <- plotcandidatemanp(pitvsdm, "PRL", "male pituitary")
e <- boxplotextint(pitvsdm, "PRL", " ")

g <- plotcandidatemanp(gonvsdm, "ESR2", "male gonads") 
h <- boxplotextint(gonvsdm, "ESR2", " ")

c <- makebargraph(DEGmanipm, "hypothalamus","DEGs w/ + LFC", 0, 3200, DEGmanipm$comparison) + 
  labs(subtitle = "male hypothalamus", x = "Comparison of manipulated group to parental controls") 
  scale_x_discrete(breaks= comparisonlevelsmanip,
                      labels= comparisonlevelsmanip,
                   drop=FALSE)

f <- makebargraph(DEGmanipm, "pituitary","DEGs w/ + LFC", 0, 3200, DEGmanipm$comparison) + 
  labs(subtitle = "male pituitary", x = "Comparison of manipulated group to parental controls") 
  scale_x_discrete(breaks= comparisonlevelsmanip,
                      labels= comparisonlevelsmanip,
                   drop=FALSE)

i <- makebargraph(DEGmanipm, "gonad","DEGs w/ + LFC", 0, 3200, DEGmanipm$comparison) + 
  theme(strip.text = element_blank()) + 
  labs(subtitle = "male gonad", x = "Comparison of manipulated group to parental controls")  + 
  scale_x_discrete(breaks= comparisonlevelsmanip,
                      labels= comparisonlevelsmanip,
                   drop=FALSE)

supplfig3 <- plot_grid(c, a,b,f,d,e,i,g,h, ncol = 3, rel_widths = c(2,1.8,1), 
                  labels = c("A", "", "",
                             "B", "", "",
                             "C", "", ""), label_size = 8, hjust = 0)
supplfig3

```
## write files

@@

pdf(file="../figures/supplfig-3-1.pdf", width=5, height=5)
plot(supplfig3)
dev.off()

write.csv(candidateDEGS, "../results/table2.csv", row.names = F)
